---
title: "p8105_hw5_yc4588"
author: "Yining Cao"
date: "2024-11-10"
output:  
  github_document:
    toc: true
    toc_depth: 2
---

### Load necessary packages
```{r setup, include=FALSE}
library(broom)
library(knitr)
library(readr)
library(tidyverse)
library(dplyr)
library(purrr)
```

## Problem 1ï¼šBirthday Simulation
```{r}
set.seed(123)

shared_birthday =  function(group_size) {
  birthdays =  sample(1:365, group_size, replace = TRUE)
  return(length(unique(birthdays)) < group_size)
}

# Run simulations and compute probabilities
birthday_simulation <- function(runs = 10000, max_group_size = 50) {
  probabilities =  numeric(max_group_size - 1)
  
  for (group_size in 2:max_group_size) {
    results =  replicate(runs, shared_birthday(group_size))
    probabilities[group_size - 1] =  mean(results)
  }
  
  return(probabilities)
}

probabilities =  birthday_simulation()

group_sizes = 2:50
plot(group_sizes, probabilities, type = "o", col = "pink", pch = 16, 
     main = "Probability of At Least Two People Sharing a Birthday",
     xlab = "Group Size", ylab = "Probability of Shared Birthday")
abline(h = 0.5, col = "skyblue", lty = 2)
legend("bottomright", legend = c("Simulation", "50% Probability"), 
       col = c("pink", "skyblue"), lty = c(1, 2), pch = c(16, NA))

```

- The function and figure illustrate the probability that at least two people in a room share a birthday, with the number of people ranging from 2 to 50. As the number of individuals in the room increases, the likelihood of a shared birthday gradually rises. By the time there are 50 people, the probability approaches 100%.

## Problem 2 : Explore power in a one-sample t-test
### Conduct a simulation to explore power in a one-sample t-test
```{r}
set.seed(123)

power_function = function(mu) {
  n = 30
  sigma = 5
  sample_data = rnorm(n, mean = mu, sd = sigma)
  test_result = t.test(sample_data, mu = 0) |>
    broom::tidy() |>
    select(estimate, p.value)
    
  return(test_result)
}

mul_mu =
  expand_grid(
  mu = 0:6,
  iter = 1:5000
) |>
  mutate(result = map(mu, power_function)) |>
  unnest(result)

power_results <- mul_mu |>
  group_by(mu) |>
  summarize(power = mean(p.value < 0.05))

ggplot(power_results, aes(x = mu, y = power)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Power of the t-test vs True Mean (mu)",
    x = "True Mean (mu)",
    y = "Power"
  ) +
  theme_minimal()
```

- The resulting plot illustrates how power increases as the true mean (Î¼) deviates further from 0, showing a stronger ability to detect larger effect sizes. When the true mean approaches 5, the power of the t-test is nearly 100%, indicating a high probability of correctly rejecting the null hypothesis.

### Plot showing that  the average estimate of ðœ‡Ì‚ vs. the true ðœ‡in all samples and in samples where the null hypothesis is rejected
```{r}
average_estimates <- mul_mu |>
  group_by(mu) |>
  summarize(
    avg_mu_hat_all = mean(estimate),
    avg_mu_hat_rejected = mean(estimate[p.value < 0.05], na.rm = TRUE)
  )

average_estimates_long <- average_estimates |>
  pivot_longer(
    cols = c(avg_mu_hat_all, avg_mu_hat_rejected),
    names_to = "type",
    values_to = "avg_mu_hat"
  ) |>
  mutate(type = recode(type,
                       "avg_mu_hat_all" = "All Samples",
                       "avg_mu_hat_rejected" = "Rejected Nulls"))

# Create plot with legend
ggplot(average_estimates_long, aes(x = mu, y = avg_mu_hat, color = type)) +
  geom_line(linewidth = 1) +
  geom_point() +
  labs(
    title = "Average Estimate of Î¼Ì‚ vs True Î¼",
    x = "True Î¼",
    y = "Average Î¼Ì‚",
    color = "Sample Type"
  ) +
  theme_minimal()
```

- No, they are not approximately equal. When the null hypothesis is rejected, the average estimate of Î¼ tends to be higher than the true value of Î¼ for smaller values of Î¼, but it approaches the true Î¼ as Î¼ increases. This happens due to selection bias: rejection of the null occurs only when the sample mean is unusually high, leading to an overestimate of Î¼ in cases where the true Î¼ is small.

## Problem 3: Analysis of Homicide Data
### Import the dataset
```{r}
homicide_data = read_csv("homicide-data.csv")
```
- The dataset contains `r nrow(homicide_data)` rows and `r ncol(homicide_data)` columns.
- Variables in the dataset include: `r names(homicide_data) |> paste(collapse = ", ")`.

### To obtain the total number of homicides and the number of unsolved homicides.
```{r}
homicide_data <- homicide_data |>
  mutate(state = replace(state, state == "wI", "WI"),
         city_state = str_c(city, state, sep = ", "),
         resolution = case_when(
           disposition == "Closed without arrest" ~ "unsolved",
           disposition == "Open/No arrest" ~ "unsolved",
           disposition == "Closed by arrest" ~ "solved")) |>
  relocate(city_state, resolution) |> 
  filter(city_state != "Tulsa, AL") 
```
- "Tulsa, AL" is removed from the dataset because it may be a typo.

### To estimate the proportion of homicides that are unsolved
```{r}
homicide_df = 
  homicide_data |> 
  group_by(city_state) |> 
  summarize(
    total = n(),
    sum_unsolved = sum(resolution == "unsolved")
  )

knitr::kable(homicide_df)
```

### Run a prop.test for the city of Balitmore, MD
```{r}
prop_test = prop.test(
  homicide_df |> filter(city_state == "Baltimore, MD") |> pull(sum_unsolved), 
  homicide_df |> filter(city_state == "Baltimore, MD") |> pull(total)) |>
  broom::tidy()
prop_test

ep_ci = prop_test |>
  select(estimate, lower_CI = conf.low, upper_CI = conf.high)
knitr::kable(ep_ci, digits = 3)

```

### To estimate the proportion of homicides that are unsolved in each city

```{r}
prop_test_results = homicide_df |>
  mutate(
    prop_tests = map2(sum_unsolved, total, \(x, y) prop.test(x = x, n = y)),
    prop_tests_tidy = map(prop_tests, broom::tidy)) |>
  select(-prop_tests) |> 
  unnest(prop_tests_tidy) |> 
  select(city_state, estimate, conf.low, conf.high) |>
  mutate(city_state = fct_reorder(city_state, estimate))

knitr::kable(prop_test_results, digits = 3)
```

### Plot showing the estimate (and CI) of the proportion of unsolved homicides in each city
```{r}
prop_test_results |> 
  ggplot(aes(x = city_state, y = estimate)) +
  geom_point(color = "skyblue", size = 1) +
  coord_flip() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), color = "red") +
  labs(title = "Estimated Proportion of Unsolved Homicides in Each City",
       x = "City, State",
       y = "Estimated Proportion of Unsolved Homicides") +
  theme_minimal()
```

